---
title: "AnnotationHub and GRanges Usecase"
author: "Sharonhe"
date: "November 2, 2016"
output: html_document
---

--
Dependencies
--

```{r}
library(GenomicRanges)
library(rtracklayer)
library(AnnotationHub)
```

---
Project 1: H3K4 trimethylation in gene promoters.
---

--
Strategy
--

1. Getting the ENCODE histone data using AnnotationHub.
2. Getting promoters using AnnotationHub.
3. Comparing the histone data and promoters using findOverlaps in GenomicRanges.

--
Getting the ENCODE histone data using AnnotationHub
--

```{r}
ah <- AnnotationHub()
ah <- subset(ah, species == "Homo sapiens")
qhs <- query(ah, "H3K4me3")
qhs <- query(qhs, "Gm12878")
qhs

```

Let's take a close look at the data

```{r}
qhs$title
qhs$dataprovider
```

Retrieve the ENCODE narrowpeak data

```[r]
gr1 <- subset(qhs, title == "wgEncodeUwHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz")[[1]]
gr1
```
Retrieve the Broad narrowpeak data

```{r}
gr2 <- subset(qhs, title == "E116-H3K4me3.narrowPeak.gz")[[1]]
gr2
```

Look at the distribution of peak widths:

```{r}
summary(width(gr1))
table(width(gr1))
summary(width(gr2))
table(width(gr2))
```

In this project, we will stick with gr1 (from ENCODE).

--
Getting promoters using AnnotationHub
--

Next we need to get some promoter coordinates. RefSeq is a highly curated collection of genes.

```{r}
qhs <- query(ah, "RefSeq")
qhs
qhs$genome
```

This looks like the same resources, but in different genome builds.

so we know which one to get:

```{r}
refseq <- qhs[qhs$genome == "hg19" & qhs$title == "RefSeq Genes"]
refseq
refseq <- refseq[[1]] ## Download the referene genome
refseq
```

```{r}
table(table(refseq$name)) ## look at the number of isoforms per gene name
```

Getting the promoters in Refseq

```{r}
promoters <- promoters(refseq)
table(width(promoters)) ## The default in this function is to use 2kb upstream and 200bp downstream of the start site.
```

--
Comparing the histone data and promoters using findOverlaps in GenomicRanges.
--

```{r}
ov <- findOverlaps(promoters, gr1)
ov
```

Compute how many percent of the peaks are in a promoter:

```{r}
length(unique(queryHits(ov))) / length(gr1)
```

Compute how many percent of promoters have a peak in them:

```{r}
length(unique(subjectHits(ov))) / length(promoters)
```

Compute a small 2x2 matrix for which bases are in promoters and/or peaks:

```{r}
prom <- reduce(promoters, ignore.strand = TRUE)
peaks <- reduce(gr1)
both <- intersect(prom, peaks)
only.prom <- setdiff(prom, both)
only.peaks <- setdiff(peaks, both)
overlapMat <- matrix(0,, ncol = 2, nrow = 2)
colnames(overlapMat) <- c("in.peaks", "out.peaks")
rownames(overlapMat) <- c("in.promoters", "out.promoter")
overlapMat[1,1] <- sum(width(both))
overlapMat[1,2] <- sum(width(only.prom))
overlapMat[2,1] <- sum(width(only.peaks))
overlapMat[2,2] <- 3*10^9 - sum(overlapMat)
round(overlapMat / 10^6, 2)
```

Compute an odds-ratio for the 2x2 table:

```{r}
oddsRatio <- overlapMat[1,1] * overlapMat[2,2] / (overlapMat[2,1] * overlapMat[1,2])
oddsRatio
```

The result shows that this odds ratio is bigger than 1, which means that it shows an enrichment of peaks in promoters.

---
Project 2: CpG islands in Human genome
---

# Use the AnnotationHub package to obtain data on "CpG Islands" in the human genome.

```{r}
ah <- AnnotationHub()
ah <- subset(ah, species == "Homo sapiens")
qhs <- query(ah, "CpG")
qhs$title
qhs$dataprovider
qhs$genome
ah_human_CpG_data <- qhs[["AH5086"]]
```

# summary info about CpG island dataset

```{r}
summary(width(ah_human_CpG_data))
seqinfo(ah_human_CpG_data)
seqlevels(ah_human_CpG_data)
gaps(ah_human_CpG_data)
```

# reduce data
```{r}
ah_human_CpG_reduce <- reduce(ah_human_CpG_data)
ah_human_CpG_reduce
```

# Calculate the islands exists on the autosomes

```{r}
autosome <- c(paste("chr", 1:22, sep=""))
split_data_by_chr <- split(ah_human_CpG_reduce, seqnames(ah_human_CpG_reduce))
autosome_CpG_data <- split_data_by_chr[autosome]
seqlevels(autosome_CpG_data)

# CpG Islands on autosome
unlist(autosome_CpG_data)

# CpG Islands on chr4
autosome_CpG_data[4]
```

---
Project 3: H3K4me3 histone modification for the H1 cell line
---

